\documentclass[journal]{IEEEtran}

\usepackage{cite}
\usepackage[]{algorithm2e}
\usepackage{color}
\usepackage{multirow}
\usepackage{subfigure}
\usepackage[justification=centering]{caption}
\usepackage{color}
\usepackage{epstopdf}
\usepackage{graphicx}
\usepackage{float} 


\usepackage[cmex10]{amsmath}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[]{algorithm2e}
\usepackage{enumitem} 



\hyphenation{op-tical net-works semi-conduc-tor}


\def\MK{\color{blue} {\bf Matthieu:} }



\begin{document}


\title{A Structured Projective Non Negative Matrix Factorization for Harmonic/Percussive Source Separation}

\author{Cl\'{e}ment~Laroche,
        Matthieu~Kowalski,
        H\'{e}l\`{e}ne~Papadopoulos,~\IEEEmembership{Member,~IEEE,}
        and Ga\"{el}~Richard,~\IEEEmembership{Senior~Member,~IEEE}% <-this % stops a space
        
}



\maketitle


\begin{abstract}
In this paper, we propose a new unconstrained nonnegative matrix factorization method designed to use the multilayer structure of audio signals in order to improve the quality of source separation. Audio signals (and music signals in particular) can be decomposed in two distinct parts: a tonal component which is sparse in frequency and temporally stable and a transient component composed of short term broadband sounds. Our method decomposes the audio signal in sparse orthogonal components that are well suited for representing the tonal component, while the transient part is represented by a regular nonnegative matrix factorization decomposition. Experiments on a large database of real music data show that such decomposition is suitable for audio source separation. Tested in comparison with three state-of-the-art harmonic/percussive decomposition algorithms, the proposed method shows competitive performances.
\end{abstract}


\begin{IEEEkeywords}
nonnegative matrix factorization, projective nonnegative matrix factorization, audio source separation, harmonic/percussive decomposition.
\end{IEEEkeywords}


\input{introduction}

\input{background}

\input{method}

\input{validation}

\input{benchmark}

\input{conclusion}

\input{appendix}


\section*{Acknowledgment}


The authors would like to thank...


\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
\bibliography{reference}


\begin{IEEEbiography}{Michael Shell}
Biography text here.
\end{IEEEbiography}

% if you will not have a photo at all:
\begin{IEEEbiographynophoto}{John Doe}
Biography text here.
\end{IEEEbiographynophoto}

% insert where needed to balance the two columns on the last page with
% biographies
%\newpage

\begin{IEEEbiographynophoto}{Jane Doe}
Biography text here.
\end{IEEEbiographynophoto}


\end{document}
